---
title: "NES-LTER IFCB Transect Winter 2018"
author: "Stace Beaulieu"
date: "15 April 2020"
output: html_document
---

## R Markdown Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# clear workspace for local development
rm(list = ls())

# set environment timezone to UTC
Sys.setenv(TZ = "UTC")

# assign relative path to directory
dir <- "/Users/sbeaulieu/Desktop/github/WHOIGit/nes-lter-ifcb-transect-winter-2018/"
# set as working directory
setwd(dir)

# define source for functions developed for the EDI packaging workflow
source("edi-utilities.R")

# install necessary libraries
# install.packages("devtools")
# install_github("EDIorg/EMLassemblyline")

# define R packages to require
libs <- c("tidyverse", "readxl", "lubridate", "devtools", "EMLassemblyline", "EML", "maps", "xml2")
# load libraries
lapply(libs, require, character.only = TRUE)
```

## Load data tables level1b and level2

Ultimately we will add columns and revise column headers here
This is not necessary for test csv files

```{r}

# note test csv files already have the columns and column headers for EDI product
level1b <- read.csv("project_folder/level1b_test.csv")
level2 <- read.csv("project_folder/level2_test.csv")

```


## QA: Map Sampling Locations

Call the map_locs function from edi-utility.R to map the sampling locations. Perform a visual check.

```{r}

# Map Check
# Note Package `maps` required for `map_data`

# level 1b
map_locs(df = level1b, xvar = "decimalLongitude", yvar = "decimalLatitude",
         region = "transect", colorvar = NULL)

# level 2
map_locs(df = level2, xvar = "longitude", yvar = "latitude",
         region = "transect", colorvar = NULL)

```


## EML Assembly: nes-lter-ifcb-transect-winter-2018

This chunk outputs the final xml file for EDI through the following steps:

Step 1: Populating EML Assembly Line templates with metadata
Step 2: Calculating the geospatial and temporal coverage 
Step 3: Making the XML file 
Step 4: Inserting a custom NES-LTER parent project node 

```{r}

# define input for EML assembly
metadata <- "IFCB-EN608-metadata"  # xlsx file
project_folder <- "project_folder/"
level1b_colhdr <- "level1b_colhdr" # metadata xlsx file
level2_colhdr <- "level2_colhdr"   # metadata xlsx file
edi_data <- c("level1b_test", "level2_test")  # data files
file_descriptions <- c("Individual-level (Level 1b) file that provides a manual classification and size data for each imaged Region of Interest (ROI)", "Sample-level (Level 2) file that summarizes the individual-level (Level 1b) file.")
pkg_id <- "knb-lter-nes.7.1"

# Make EML Templates
# copy abstract, additional_info, and methods txt files into project_folder
# or else this next step will produce empty files
# Because this package will have two tables,
# this first xlsx does not have attributes for either

xlsx_to_template(metadata.path = paste0(dir, metadata),
                 output.path = paste0(dir, project_folder),
                 edi.filename = NULL, 
                 rights = "CCBY")
# level1b attributes needs separate xlxs
xlsx_to_template(metadata.path = paste0(dir, project_folder, level1b_colhdr), 
                 output.path = paste0(dir, project_folder),
                 edi.filename = level1b_colhdr, 
                 rights = "CCBY")
# level2 attributes needs separate xlxs
xlsx_to_template(metadata.path = paste0(dir, project_folder, level2_colhdr), 
                 output.path = paste0(dir, project_folder),
                 edi.filename = level2_colhdr, 
                 rights = "CCBY")

# Data Coverage
# use level2
# isolate date and geospatial columns for input
date_col <- as.Date(level2$date)
lat_col <- level2$latitude
lon_col <- level2$longitude
# run function to determine geospatial and temporal coverage
coverage <- data_coverage(dates = date_col, lat = lat_col, lon = lon_col)

# Make EML
make_eml(path = paste0(dir, project_folder),
         dataset.title = "Abundance and biovolume of taxonomically-resolved phytoplankton imaged continuously underway with an Imaging FlowCytobot along the NES-LTER Transect in winter 2018",
         data.table = paste0(edi_data, ".csv"),
         data.table.description = file_descriptions,
         temporal.coverage = c(coverage$startdate, coverage$enddate),
         geographic.description = "NES-LTER Transect",
         geographic.coordinates = c(coverage$North, coverage$East, coverage$South, coverage$West),
         maintenance.description = "completed",
         user.id = "NES",
         user.domain = "LTER",
         package.id = pkg_id)

# Insert Custom Project Node
project_insert(edi_pkg = pkg_id, 
               xml.path = paste0(dir, project_folder))
```

## Manually check a couple of ROIs in the data product against the IFCB Dashboard
