---
title: "NES-LTER IFCB Transect Winter 2018"
author: "Stace Beaulieu"
date: "08 May 2020"
output: html_document
---

## R Markdown Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# clear workspace for local development
rm(list = ls())

# set environment timezone to UTC
Sys.setenv(TZ = "UTC")

# assign relative path to directory
dir <- "/Users/sbeaulieu/Desktop/github/WHOIGit/nes-lter-ifcb-transect-winter-2018/"
# set as working directory
setwd(dir)

# define source for functions developed for the EDI packaging workflow
source("edi-utilities.R")

# install necessary libraries
# install.packages("devtools")
# install_github("EDIorg/EMLassemblyline")

# define R packages to require
libs <- c("tidyverse", "readxl", "lubridate", "devtools", "EMLassemblyline", "EML", "maps", "xml2")
# load libraries
lapply(libs, require, character.only = TRUE)

# # install emlassemblyline fix_31b to try the semantic annotation
# remotes::install_github("EDIorg/EMLassemblyline", ref = "fix_31b")
# library(EMLassemblyline)

```

## Load data tables level1b and level2 revise column headers and add columns here for OBIS

```{r}

# level1b <- read_csv("https://github.com/klqi/EDI-NES-LTER-2019/raw/f040f31c290ff40baf9a48c9ce2dd57b46336900/auto_join/level_1b_manual.csv")  # does not account for 7 ROIs with features zero
level1b <- read_csv("https://github.com/klqi/EDI-NES-LTER-2019/raw/d61d55de36137e94bfafa94c080849f4554e5df6/auto_join/level_1b_manual.csv") # accounts for 7 ROIs with features zero
# view summary stats to ensure no features zero
summary(level1b)
# check for 7 NAs
checkNAlevel1b <- level1b %>% filter(is.na(Area))

# level1b needs an occurrenceID
# make a copy of the column
level1b$occurrenceID <- level1b$associatedMedia
# strip prefix http://ifcb-data.whoi.edu/
level1b$occurrenceID <- gsub("http://ifcb-data.whoi.edu/", "", level1b$occurrenceID)
# strip suffix .html
level1b$occurrenceID <- gsub("\\..*", "", level1b$occurrenceID)

# level1b needs an eventID to be able to join with level2
# example: use http://ifcb-data.whoi.edu/NESLTER_transect/D20180201T103729_IFCB102_00002.html
# to generate NESLTER_transect/D20180201T103729_IFCB102
# make a copy of the column
level1b$eventID <- level1b$occurrenceID
# strip suffix after IFCB102
level1b$eventID <- gsub("(IFCB102).*", "\\1", level1b$eventID)


# need level 2 data to be able to add columns required for OBIS occurrence table
# eventDate, decimalLatitude, decimalLongitude, depth
# level2 <- read_csv("https://github.com/klqi/EDI-NES-LTER-2019/raw/f040f31c290ff40baf9a48c9ce2dd57b46336900/plot1/level_2_manual.csv") # does not account for 7 ROIs with features zero
level2 <- read_csv("https://github.com/klqi/EDI-NES-LTER-2019/raw/d61d55de36137e94bfafa94c080849f4554e5df6/plot1/level_2_manual.csv") # changed zero features to NaN commit d61d55d

# make sure sum of level2 abundance equals the number of ROIs in level1b
sum(level2$Abundance)
# make sure sum of level2 biovolume equals the sum of level1b biovolume
sum(level2$Biovolume)
sum(level1b$Biovolume, na.rm=TRUE) 
# view summary stats, confirm range volume_imaged
summary(level2)

# rename level2 columns to eventDate, decimalLatitude, decimalLongitude, higherClassification_group
level2 <- rename(level2, eventDate = date)
level2 <- rename(level2, decimalLatitude = gps_furuno_latitude)
level2 <- rename(level2, decimalLongitude = gps_furuno_longitude)
level2 <- rename(level2, higherClassification_group = higherClassification_group_manual)
level2$depth = 5

# level2 needs an eventID to be able to join with level1
# make a copy of the column
level2$eventID <- level2$associatedMedia
# strip prefix http://ifcb-data.whoi.edu/
level2$eventID <- gsub("http://ifcb-data.whoi.edu/", "", level2$eventID)
# strip suffix .html
level2$eventID <- gsub("\\..*", "", level2$eventID)

# create dataframe with only the unique attributes assoc'd to eventIDs from level2
# eventID, eventDate, decimalLatitude, decimalLongitude, plus depth
uniqueeventIDs <- select(level2, eventID, eventDate, decimalLatitude, decimalLongitude, depth)
uniqueeventIDs <- distinct(uniqueeventIDs)
# unique eventIDs should be 60 for IFCB EN608 dataset
# manually inspect eventDate for temporal coverage

# full_join level1b with uniqueeventIDs from level2 to make sure no samples missing
level1bEDI <- full_join(level1b, uniqueeventIDs, by = "eventID")

# need level 1b data to have more columns required for OBIS occurrence table
level1bEDI$occurrenceStatus = "present"
level1bEDI$basisOfRecord = "HumanObservation"

# strip level1b extraneous column "X1"
level1bEDI <- select(level1bEDI, -X1)

# add column concentrationFlag to level2 esp. since IFCB triggered only on fluorescence
level2$concentrationFlag = 1  # initiate concentrationFlag with 1 likely valid

# for the following data_provider_category_HumanObservation replace concentrationFlag with zero may be invalid
# 20200414_classify_classlabel_matched_sorted_concentrationFlag.xls
# the following as reported by Brownlee et al. (2016) https://doi.org/10.3354/meps11687
# "ciliate", "Eutintinnus", "Gyrodinium", "Tintinnida", "Tintinnidium mucicola", "Favella", "Oxyrrhis", "Protoperidinium"
# the following because heterotrophs or the group includes heterotrophs
# "amoeba", "Dinophyceae", "flagellate", "foraminifera", "Nauplii", "Parvicorbicula socialis", "Radiozoa", "zooplankton"
# the following because not organisms
# "bead", detritus", "detritus clear", "fecal pellet", "other"
# the following empirical
# "cryptophyta" undercounted on EN608, "Leptocylindrus mediterraneus", "Helicostomella subulata",  "Polykrikos"
# the following at smallest or largest end of size range
# "mix", "Coscinodiscus"

undersampled <- list("amoeba", "bead", "ciliate", "Coscinodiscus", "cryptophyta", "detritus", "detritus clear", "Dinophyceae", "Eutintinnus", "Favella", "fecal pellet", "flagellate", "foraminifera", "Gyrodinium", "Helicostomella subulata", "Leptocylindrus mediterraneus", "mix","Nauplii", "other", "Oxyrrhis", "Parvicorbicula socialis", "Polykrikos", "Protoperidinium", "Radiozoa", "Tintinnida", "Tintinnidium mucicola", "zooplankton")

for (i in undersampled) {
level2 %>% mutate(concentrationFlag = replace(concentrationFlag, data_provider_category_HumanObservation == i, 0)) -> level2
}

# Level 2 round volume_imaged to 3 past decimal
level2$volume_imaged <- round(level2$volume_imaged, digits = 3)

```


## QA: Map Sampling Locations

Call the map_locs function from edi-utility.R to map the sampling locations. Perform a visual check.

```{r}

# Map Check
# Note Package `maps` required for `map_data`

# level 1b
map_locs(df = level1bEDI, xvar = "decimalLongitude", yvar = "decimalLatitude",
         region = "transect", colorvar = NULL)

# level 2
map_locs(df = level2, xvar = "decimalLongitude", yvar = "decimalLatitude",
         region = "transect", colorvar = NULL)

```

## Additional quality assurance for data values

```{r}
# these steps in addition to summary stats when reading in data tables

# level 1b all occurrenceIDs must be unique
length(unique(level1bEDI$occurrenceID))

# level 1b minFeret, sqrt area, and cube root biovolume should be < 150 micron
toowide <- level1bEDI %>% filter(minFeretDiameter > 150)
# there were 8 ROIs that with minFeretDiameter slightly greater than 150
# these were mainly Ceratium
tooarea <- level1bEDI %>% filter(sqrt(Area) > 150)
toobiovol <- level1bEDI %>% filter(Biovolume^(1/3) > 150)

# level 2 all volume_imaged must be < 5 mL
toovol <- level2 %>% filter(volume_imaged > 5)

# level 2 make sure no NAs in abundance or biovolume
abundNA <- level2 %>% filter(is.na(Abundance))
biovolNA <- level2 %>% filter(is.na(Biovolume))

```

## Column Header Organization and write files for upload to EDI
```{r}
# define headers for columns in desired order
level1bEDI_headers <- c("associatedMedia","data_provider_category_HumanObservation","scientificName_HumanObservation","scientificNameID_HumanObservation","higherClassification_group","Area","Biovolume","maxFeretDiameter","minFeretDiameter","eventDate","decimalLatitude","decimalLongitude","depth","eventID","occurrenceID","occurrenceStatus","basisOfRecord")
level2_headers <- c("associatedMedia","data_provider_category_HumanObservation","scientificName_HumanObservation","scientificNameID_HumanObservation","higherClassification_group","Abundance","Biovolume","volume_imaged","concentrationFlag","eventDate","decimalLatitude","decimalLongitude","depth","eventID")

# reorder columns as necessary
level1b_edi <- level1bEDI[, level1bEDI_headers]
level2_edi <- level2[, level2_headers]

# write files for upload to EDI
write.csv(level1b_edi, paste0(dir, "/project_folder/level1b_edi.csv"),
          row.names = FALSE)
write.csv(level2_edi, file.path(dir, "/project_folder/level2_edi.csv"),
          row.names = FALSE)

```


## EML Assembly: nes-lter-ifcb-transect-winter-2018

This chunk outputs the final xml file for EDI through the following steps:

Step 1: Populating EML Assembly Line templates with metadata
Step 2: Calculating the geospatial and temporal coverage 
Step 3: Making the XML file 
Step 4: Inserting a custom NES-LTER parent project node 

```{r}

# define input for EML assembly
metadata <- "IFCB-EN608-metadata"  # xlsx file
project_folder <- "project_folder/"
# level1b_test <- "level1b_test" # metadata xlsx file
# level2_test <- "level2_test"   # metadata xlsx file
level1b_edi <- "level1b_edi" # metadata xlsx and data csv files
level2_edi <- "level2_edi" # metadata xlsx and data csv files
# edi_data <- c("level1b_test", "level2_test")  # data files
edi_data <- c("level1b_edi", "level2_edi")  # data files
file_descriptions <- c("Individual-level (Level 1b) file that provides classification and size for each imaged Region of Interest (ROI)", "Sample-level (Level 2) file that summarizes the individual-level (Level 1b) file")
pkg_id <- "knb-lter-nes.9.1"

# Make EML Templates
# copy abstract, additional_info, and methods txt files into project_folder
# or else this next step will produce empty files
# Because this package will have two tables,
# this first xlsx does not have attributes or catvars templates

xlsx_to_template(metadata.path = paste0(dir, metadata),
                 output.path = paste0(dir, project_folder),
                 edi.filename = NULL, 
                 rights = "CCBY")
# # level1b attributes needs separate xlxs
# # do not confuse with data csv file
# xlsx_to_template(metadata.path = paste0(dir, level1b_test), 
#                  output.path = paste0(dir, project_folder),
#                  edi.filename = level1b_test, 
#                  rights = "CCBY")
xlsx_to_template(metadata.path = paste0(dir, level1b_edi), 
                 output.path = paste0(dir, project_folder),
                 edi.filename = level1b_edi, 
                 rights = "CCBY")
# # level2 attributes needs separate xlxs
# # do not confuse with data csv file
# xlsx_to_template(metadata.path = paste0(dir, level2_test), 
#                  output.path = paste0(dir, project_folder),
#                  edi.filename = level2_test, 
#                  rights = "CCBY")
xlsx_to_template(metadata.path = paste0(dir, level2_edi), 
                 output.path = paste0(dir, project_folder),
                 edi.filename = level2_edi, 
                 rights = "CCBY")

# Data Coverage
# use level2
# isolate date and geospatial columns for input
date_col <- as.Date(level2$eventDate)
lat_col <- level2$decimalLatitude
lon_col <- level2$decimalLongitude
# run function to determine geospatial and temporal coverage
coverage <- data_coverage(dates = date_col, lat = lat_col, lon = lon_col)

# Make EML
make_eml(path = paste0(dir, project_folder),
         dataset.title = "Abundance and biovolume of taxonomically-resolved phytoplankton and microzooplankton imaged continuously underway with an Imaging FlowCytobot along the NES-LTER Transect in winter 2018",
         data.table = paste0(edi_data, ".csv"),
         data.table.description = file_descriptions,
         temporal.coverage = c(coverage$startdate, coverage$enddate),
         geographic.description = "NES-LTER Transect",
         geographic.coordinates = c(coverage$North, coverage$East, coverage$South, coverage$West),
         maintenance.description = "completed",
         user.id = "NES",
         user.domain = "LTER",
         package.id = pkg_id)

# Insert Custom Project Node
project_insert(edi_pkg = pkg_id, 
               xml.path = paste0(dir, project_folder))
```

## Add semantic annotations to EML 2.2

```{r}

# install emlassemblyline fix_31b to try the semantic annotation
# I already prepared an annotations.txt file for the keyword image

remotes::install_github("EDIorg/EMLassemblyline", ref = "fix_31b")
library(EMLassemblyline)

annotate_eml(annotations = "project_folder/annotations.txt", eml.in = "project_folder/knb-lter-nes.9.1.xml", eml.out = "knb-lter-nes.9.1.xml")

# REMEMBER TO REINSTALL THE REGULAR EMLASSEMBLYLINE AFTERWARDS BECAUSE CAN'T GENERATE EML YET FROM SCRATCH WITH THIS BRANCH
# # Install and load EMLassemblyline
# remotes::install_github("EDIorg/EMLassemblyline")
# library(EMLassemblyline)

```

